Query Optimization for Dynamic Imputation

### 2017-07-10 paper/main.tex
paper: conform to camera ready style
add qcri acknowledgment
feat: start cleaning up command line interface
fig 4: fix to avoid dirty set issue
some changes for camera-ready, see slack post
paper: swap comment symbol
paper: fix overfull hbox and length problems
paper: add github url
paper: update github url
Update constants

\documentclass[preprint]{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=savetrees,firstinits=true,maxcitenames=2]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{stmaryrd}
\usepackage{qtree} % for simple tree diagrams
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{xspace}
\usepackage[all]{nowidow}
\usepackage{enumitem}
\usepackage{hyperref} % load after other packages
\usepackage{cleveref}

@@

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

@@

\newcommand{\ProjectName}{ImputeDB\xspace}
\newcommand{\sys}{\ProjectName}
\newcommand{\doms}[2]{#1 \succ #2}

@@

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color{red}\textbf{TODO:} #1}}
\newcommand{\todobox}[2]{{\color{red}\fbox{\parbox{\columnwidth}{\textbf{TODO:} \textit{#1}\\#2}}}}
\newcommand{\review}[1]{{\color{blue}#1}}

% results that can be used in any section
\newcommand{\demorows}{10175}
\newcommand{\labexrows}{9813}
\newcommand{\acsbaseresultminutes}{355 minutes}
\newcommand{\acsbaseresulthours}{just under 6 hours}
\newcommand{\acsimputedbzeroresult}{4 seconds}
\newcommand{\acsimputedboneresult}{1 second}
\newcommand{\lowxalphazero}{10} % the smallest speed up, comparing a=0 to base table imputation
\newcommand{\highxalphazero}{140} % the largest speed up, comparing a=0 to base table imputation 
\newcommand{\highxalphaone}{1400} % the largest speed up, comparing a=0 OR a=1 to base table imputation 
\newcommand{\lowsmapealphazero}{0} % the smallest SMAPE difference, comparing a=0 to base table imputation
\newcommand{\highsmapealphazero}{6} % the largest SMAPE difference, comparing a=0 to base table imputation
\newcommand{\highsmapealphaone}{20} % the largest SMAPE difference, comparing a=1 to base table imputation
%\newcommand{\highxalphaone}{16000}
\newcommand{\nullv}{NULL}

@@

\begin{abstract}
  Missing values are common in data analysis and present a usability challenge.
  Users are forced to pick between removing tuples with missing values or creating a cleaned version of their data by applying a relatively expensive imputation strategy.
  Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, performing necessary imputations on-the-fly for each query.
  This allows users to immediately explore their data, while the system picks the optimal placement of imputation operations.
  % This placement is based on the specific query issued and a user-provided parameter that trades off data quality and runtime performance.
  % We incorporate the imputation operators into the relational algebra, allowing for traditional optimizations to take place.
  We evaluate this approach on three real-world survey-based datasets.
  % from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}.
  Our experiments show that our query plans execute between \lowxalphazero{} and \highxalphazero{} times faster than first imputing the base tables.
  Furthermore, we show that the query results from on-the-fly imputation differ from the
  traditional base-table imputation approach by \lowsmapealphazero{}--\highsmapealphaone{}\%.
  Finally, we show that while dropping tuples with missing values that fail query constraints discards 6--78\% of the data, on-the-fly imputation loses only 0--21\%.
\end{abstract}

@@

\section{Acknowledgments}
The authors would like to acknowledge \AtNextCite{\defcounter{maxnames}{3}}\citeauthor{akande2015empirical} for giving us a cleaned copy of the 2012 ACS PUMS.

\bibstyle{abbrv}
\printbibliography
\balance

>>>>

\documentclass{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle,bibliography]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=savetrees,firstinits=true,maxcitenames=2]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{stmaryrd}
\usepackage{qtree} % for simple tree diagrams
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{xspace}
\usepackage[all]{nowidow}
\usepackage{enumitem}
\usepackage{hyperref} % load after other packages
\usepackage{cleveref}
\usepackage{mathtools}

@@

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\commentsymbol}[0]{$\blacktriangleright$}
\renewcommand{\algorithmiccomment}[1]{\hskip2em\commentsymbol\emph{#1}}

@@

\newcommand{\ProjectName}{ImputeDB\xspace}
\newcommand{\ProjectUrl}{\url{https://github.com/mitdbg/imputedb}}
\newcommand{\sys}{\ProjectName}
\newcommand{\doms}[2]{#1 \succ #2}

@@

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color{red}\textbf{TODO:} #1}}
\newcommand{\todobox}[2]{{\color{red}\fbox{\parbox{\columnwidth}{\textbf{TODO:} \textit{#1}\\#2}}}}
\newcommand{\review}[1]{#1}
\newcommand{\cameraready}[1]{{\color{red}#1}}

% results that can be used in any section
\newcommand{\demorows}{10175}
\newcommand{\labexrows}{9813}
\newcommand{\acsbaseresultminutes}{355 minutes}
\newcommand{\acsbaseresulthours}{just under 6 hours}
\newcommand{\acsimputedbzeroresult}{4 seconds}
\newcommand{\acsimputedboneresult}{1 second}

\newcommand{\lowxalphazero}{10} % the smallest speed up, comparing a=0 to base table imputation
\newcommand{\highxalphazero}{140} % the largest speed up, comparing a=0 to base table imputation 
\newcommand{\highxalphaone}{1400} % the largest speed up, comparing a=0 OR a=1 to base table imputation

\newcommand{\lowsmapealphazero}{0} % the smallest SMAPE difference, comparing a=0 to base table imputation
\newcommand{\highsmapealphazero}{8} % the largest SMAPE difference, comparing a=0 to base table imputation
\newcommand{\highsmapealphaone}{20} % the largest SMAPE difference, comparing a=1 to base table imputation

\newcommand{\lowcountfractionlostalphazero}{0}
\newcommand{\highcountfractionlostalphazero}{21}
\newcommand{\lowcountfractionlostalphaone}{6}
\newcommand{\highcountfractionlostalphaone}{78}

%\newcommand{\highxalphaone}{16000}
\newcommand{\nullv}{NULL}

@@

\begin{abstract}
  Missing values are common in data analysis and present a usability challenge.
  Users are forced to pick between removing tuples with missing values or creating a cleaned version of their data by applying a relatively expensive imputation strategy.
  Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, performing necessary imputations on-the-fly for each query.
  This allows users to immediately explore their data, while the system picks the optimal placement of imputation operations.
  % This placement is based on the specific query issued and a user-provided parameter that trades off data quality and runtime performance.
  % We incorporate the imputation operators into the relational algebra, allowing for traditional optimizations to take place.
  We evaluate this approach on three real-world survey-based datasets.
  % from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}.
  Our experiments show that our query plans execute between \lowxalphazero{} and \highxalphazero{} times faster than first imputing the base tables.
  Furthermore, we show that the query results from on-the-fly imputation differ from the
  traditional base-table imputation approach by \lowsmapealphazero{}--\highsmapealphaone{}\%.
  Finally, we show that while dropping tuples with missing values that fail query constraints discards \lowcountfractionlostalphaone--\highcountfractionlostalphaone\% of the data, on-the-fly imputation loses only \lowcountfractionlostalphazero--\highcountfractionlostalphazero\%.
\end{abstract}

@@

\section{Acknowledgments}
The authors would like to acknowledge \AtNextCite{\defcounter{maxnames}{3}}\citeauthor{akande2015empirical} for giving us a cleaned copy of the 2012 ACS PUMS.
This work was supported in part by the Qatar Research Computing Institute.

\emergencystretch=1em
\bibstyle{abbrv}
\printbibliography
\balance
